---
title: "1.1 Figure 2"
author: "TG"
date: "2023-06-27"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(plotrix)
library(sjPlot)

color_op6 <- c("#20A39E","#23001E","#FFBA49","#EF5B5B")

save_path<-"G:\\My Drive\\PHD\\chapter 1 strom effect\\R_storm_ms\\plots\\Fig 2\\Main plot"
model_path<-"G:\\My Drive\\PHD\\chapter 1 strom effect\\R_storm_ms\\models\\Table 2\\main table"
```


In this script ill concentrate all the models and plot constructing Fig 2

# Data prepartion

```{r}
load("fish_full_data.rdata")
select <- dplyr::select
```

## general steps:

1. keep only 2018 - 2021

```{r}

fish_prep<-fish_full_data %>% filter(Year>2017)

```


2. remove irregular sites

```{r}
knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

fish_prep<-fish_prep %>% filter(!Knoll %in% knolls_to_remove) %>% filter(Site !="North beach")
```


3. keep one observer from each knoll

```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(fish_prep$survey_id)) {
  
  data <- fish_prep %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

fish_prep <- bind_rows(one_observer)

rm(data,filter_data,one_observer)

# make sure i have the right knolls number


knoll_site_year <-fish_prep %>% group_by(Site,year_month) %>% summarise("knoll_number"= n_distinct(knoll_id))

```

4. keep only the resident fish

```{r}

fish_prep <- fish_prep %>% filter(Status != "T")

```


5. convert year_month to factor

```{r}

fish_prep$year_month<-as.factor(fish_prep$year_month)
```

6. create the data for species richness, abundance and shannon model

General stepes:

1. filter accourding to confidance level (richness and eveness; confidance = 0, abundance; confidance = 2)

2. filter schooling fish for evenness calculations

3. group by the right variable (number of species/ind/effective number)


### richness

```{r}

richness <- fish_prep %>% filter(Confidence < 1)

richness<- richness %>%
  group_by(survey_id) %>%
  mutate(Species_number = n_distinct(Species)) %>%
  select(-Species,-Species_2015,-Family,-Abundance,-Length,-Confidence,-Status,-a,-b) %>%
  ungroup() %>%
  distinct(.keep_all = T)

```



### abundance

```{r}
abundance <- fish_prep %>%
  group_by(survey_id,Site,Knoll,year_month,Max_depth,Min_depth,Mean_Complexity,
           Surface_Area,Algae,coral_cover) %>%
    summarise("ind"=sum(Abundance))

which(!abundance$survey_id %in% richness$survey_id) # need to be 0
```

### Shannon

first - filter and convert to wide data format


```{r}

school_species <- c("Chromis viridis",
                    "Pseudanthias squamipinnis",
                    "Neopomacentrus miryae",
                    "Atherinomorus lacunosus",
                    "Spratelloides gracilis") 

shannon <- fish_prep %>% filter(! Species %in% school_species) %>% filter(Confidence < 1)


shannon <- shannon %>%
  group_by(Site,Year,year_month,Knoll,DD_X,DD_Y,Surface_Area,Max_depth,Date,survey_id,Species) %>%
  summarise("aboundance"=sum(Abundance))


shannon<-spread(shannon,Species,aboundance,fill = 0)

which(!abundance$survey_id %in% shannon$survey_id) #need to be 1,transect "NR_4_2020-06-15" had only schooling fish
```
calculate the shannon effective number

```{r}
shannon$shannon_div<-diversity(shannon[,11:ncol(shannon)],index="shannon")

shannon<-shannon %>%
  relocate(shannon_div,.after=survey_id) %>% 
  mutate("shannon_effective"=exp(shannon_div),.after=shannon_div)
  


```

# Models

## Richness model

```{r}
richness_model<-lm(formula = log(Species_number) ~ log(Surface_Area)*year_month ,data = richness)

summary(richness_model)
r.squaredGLMM(richness_model)

plot(richness_model)

richness_model_gg = ggeffects::ggpredict(richness_model,terms = c('Surface_Area[1:66 by=1]','year_month'))

tab_model(richness_model,file = paste0(model_path,"\\richness_model.xls"))


ggplot()+
  theme_classic()+
  geom_line(data = richness_model_gg,aes(x= log(x),y = log(predicted),color= group),size=1.5)+ 
  geom_ribbon(data = richness_model_gg,aes(x=log(x),ymin=log(conf.low),ymax=log(conf.high),fill=group),alpha=0.2)+
  ylab("Log(Richness)")+xlim(1,4)+ylim(-1,4)+
  xlab(bquote('Log(Surface area '~(m^2)))+
    theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
        scale_color_manual(values=color_op6,
                           name = "Year",
                           labels = c("Fall 2018","Spring 2020*", "Fall 2020" ,
                                      "Spring 2021"))+
        scale_fill_manual(values=color_op6,
                           name = "Year",
                           labels = c("Fall 2018","Spring 2020*", "Fall 2020" ,
                                      "Spring 2021"))


#ggsave("Area_richness_18.png" , width = 12, height = 5, dpi = 300,path = save_path)
#ggsave("Area_richness_18.pdf" , width = 12, height = 5, dpi = 300,path = save_path)

```

## Abundance model


```{r}
abundance_model<-lm(formula = log(ind) ~ log(Surface_Area)*year_month,data = abundance)
summary(abundance_model)

tab_model(abundance_model,file = paste0(model_path,"\\abundance_model.xls"))

plot(abundance_model)


abundance_model_gg=ggeffects::ggpredict(abundance_model,terms = c('Surface_Area[1:66 by=1]','year_month'))

ggplot()+
  theme_classic()+
  geom_line(data = abundance_model_gg,aes(x= log(x),y = log(predicted),color= group),size=1.5)+ 
  geom_ribbon(data = abundance_model_gg,aes(x=log(x),ymin=log(conf.low),ymax=log(conf.high),fill=group),alpha=0.2)+
  ylab("Log(Abundabce)")+xlim(1,4)+
  xlab(bquote('Log(Surface area'~(m^2)))+
    theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
        scale_color_manual(values=color_op6,
                           name = "Year",
                           labels = c("Fall 2018","Spring 2020*", "Fall 2020" ,
                                      "Spring 2021"))+
        scale_fill_manual(values=color_op6,
                           name = "Year",
                           labels = c("Fall 2018","Spring 2020*", "Fall 2020",
                                      "Spring 2021"))


#ggsave("Area_abundance_18.png" , width = 12, height = 5, dpi = 300,path = save_path)
#ggsave("Area_abundance_18.pdf" , width = 12, height = 5, dpi = 300,path = save_path)


```

## Shannon model

```{r}

shannon_model<-lm(formula = log(shannon_effective) ~ log(Surface_Area)*year_month,data = shannon)
summary(shannon_model)

tab_model(shannon_model,file = paste0(model_path,"\\shannon_model.xls"))


shannon_model_gg=ggeffects::ggpredict(shannon_model,terms = c('Surface_Area[1:66 by=1]','year_month'))

ggplot()+
  theme_classic()+
  geom_line(data = shannon_model_gg,aes(x= log(x),y = log(predicted),color= group),size=1.5)+ 
  geom_ribbon(data = shannon_model_gg,aes(x=log(x),ymin=log(conf.low),ymax=log(conf.high),fill=group),alpha=0.2)+
  ylab("Log(Effective number of species)")+xlim(1,4)+
  xlab(bquote('Log(Surface area'~(m^2)))+
    theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))+
        scale_color_manual(values=color_op6,
                           name = "Year",
                           labels = c("Fall 2018","Spring 2020*", "Fall 2020" ,
                                      "Spring 2021"))+
        scale_fill_manual(values=color_op6,
                           name = "Year",
                           labels = c("Fall 2018","Spring 2020*", "Fall 2020" ,
                                      "Spring 2021"))

#ggsave("Area_shannon_18.png" , width = 12, height = 5, dpi = 300,path = save_path)
#ggsave("Area_shannon_18.pdf" , width = 12, height = 5, dpi = 300,path = save_path)

```

# bar plots

## richness

create wide data


```{r}
richness_rarefied <- fish_prep %>% filter(Confidence < 1) %>% 
  group_by(Site,Year,year_month,Knoll,Surface_Area,Max_depth,Date,survey_id,Species) %>%
  summarise("abundance"=sum(Abundance))


richness_rarefied<-spread(richness_rarefied,Species,abundance,fill = 0)
```


run rarefaction function (sample based)

```{r}


list_richness_rare<-list()
loop<-1

for (i in unique(richness_rarefied$year_month)) {
  
  year<- richness_rarefied %>% filter(year_month == i)
  
  year_richness<-Rarefy::rare_alpha(year[9:ncol(year)],
                          method="hill",
                          q=0,
                          random=50,
                          mean = F)
  
  year_richness$Year<-rep(i)
  year_richness$sample<-1:nrow(year_richness)
  
  
  list_richness_rare[[loop]]<-year_richness
  loop<-loop+1
  
  
}

year_richness<-bind_rows(list_richness_rare)


year_richness<-year_richness %>% filter(sample == 50)


ggplot(data = year_richness,aes(x=Year,y=Rarefaction,fill = Year))+
geom_bar(stat="identity")+geom_errorbar(aes(ymin=IC_low,ymax= IC_up),width =0.1)+
  scale_fill_manual(values = color_op6)+
  xlab("")+ 
  ylab('Richness')+
  theme_classic()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20))+
    scale_x_discrete(labels = c("Fall 2018","Spring 2020",
                                      "Fall 2020" ,
                                      "Spring 2021"))


#ggsave("rare_richness_bar_18.png" , width = 7, height = 5, dpi = 300,path = save_path)
#ggsave("rare_richness_bar_18.pdf" , width = 7, height = 5, dpi = 300,path = save_path)

```


## abundance

*** still didnt change


```{r}
ind_per_m<- abundance %>%  group_by(Site,year_month,Knoll,Surface_Area) %>% summarise("inds"=sum(ind)) %>% mutate("inds_m"=inds/Surface_Area)


ind_per_m <- ind_per_m %>%
  group_by(year_month) %>%
  summarise("mean_density"=mean(inds_m),                                  "se"= std.error(inds_m),
            "n"=n())


ggplot(data = ind_per_m,aes(x=year_month,y=mean_density,fill = year_month))+
geom_bar(stat="identity")+geom_errorbar(aes(ymin=mean_density-se,ymax=mean_density+se),width =0.1)+
  geom_text(aes(label = n,y=1), size = 7,color= "white")+  
  xlab("Year")+ 
  ylab(bquote('Mean fish density '(Individuals/m^2)))+
  theme_classic()+
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20))+
  scale_fill_manual(values= color_op6,
                           name = "Year",
                           labels = c("Fall 2018","Spring 2020",
                                      "Fall 2020" ,
                                      "Spring 2021"))+
  scale_x_discrete(labels = c("Fall 2018","Spring 2020",
                                      "Fall 2020" ,
                                      "Spring 2021"))


#ggsave("abundance_bar_18.png" , width = 7, height = 5, dpi = 300,path = save_path)
#ggsave("abundance_bar_18.pdf" , width = 7, height = 5, dpi = 300,path = save_path)
```



## shannon

sample based rarefaction for effective number of species by shannon index

```{r}


shannon_rare<-list()
loop<-1



for (i in unique(shannon$year_month)) {
  
  year<- shannon %>% filter(year_month == i)
  
  year_richness_shannon<-Rarefy::rare_alpha(year[13:ncol(year)],
                          method="hill",
                          q=1,
                          random=50,
                          mean = F)
  
  year_richness_shannon$Year<-rep(i)
  year_richness_shannon$sample<-1:nrow(year_richness_shannon)
  
  
  shannon_rare[[loop]]<-year_richness_shannon
  loop<-loop+1
  
  
}

richness_shannon<-bind_rows(shannon_rare)
```


```{r}
richness_shannon<-richness_shannon %>% filter(sample == 50)


ggplot(data = richness_shannon,aes(x=Year,y=Rarefaction,fill = Year))+
geom_bar(stat="identity")+geom_errorbar(aes(ymin=IC_low,ymax= IC_up),width =0.1)+
  scale_fill_manual(values = color_op6)+
  xlab("")+ 
  ylab('Effective number of species')+
  theme_classic()+ 
    theme(axis.text.x = element_text(size=20),
        plot.title = element_text(size=24),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        legend.position = "none",
        strip.text.x = element_text(size = 20))+
    scale_x_discrete(labels = c("Fall 2018","Spring 2020",
                                      "Fall 2020" ,
                                      "Spring 2021"))

#ggsave("rare_shannon_bar_18.png" , width = 7, height = 5, dpi = 300,path = save_path)
#ggsave("rare_shannon_bar_18.pdf" , width = 7, height = 5, dpi = 300,path = save_path)



```

