---
title: "2.1 log ratio plots"
author: "TG"
date: "2023-06-27"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r message=FALSE, warning=FALSE}

library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(Rarefy)
library(rareNMtests)
library(zetadiv)
library(FuzzyQ)
library(SingleCaseES)

```



color scales definitions 

```{r}
color_op6 <- c("#20A39E","#23001E","#FFBA49","#EF5B5B")
rank_filter <-51
plot_path<-"G:\\My Drive\\PHD\\chapter 1 strom effect\\R_storm_ms\\plots\\Log raito"
```


# Data prepartion

```{r}
load("fish_full_data.rdata")
select <- dplyr::select

fish_prep<-fish_full_data
```


2. keep only data from 2018 
```{r}

fish_prep<- fish_prep %>% filter(Year > 2017)

```



3. remove irregular sites


```{r}

knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

fish_prep<-fish_prep %>% filter(!Knoll %in% knolls_to_remove) %>% 
  filter(Site !="North beach")

```

4. keep only sure identifications

filter confidence = 0


```{r}

fish_prep<-fish_prep %>% filter(Confidence <1)

```

5. keep one observer from each knoll


```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(fish_prep$survey_id)) {
  
  data <- fish_prep %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
  
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

fish_prep <- bind_rows(one_observer)

rm(data,filter_data,one_observer)

# make sure i have the right knolls number


knoll_site_year <-fish_prep %>% group_by(Site,year_month) %>% summarise("knoll_number"= n_distinct(knoll_id))

```


6. convert year_month to factor

```{r}

fish_prep$year_month<-as.factor(fish_prep$year_month)

```



7. keep only the knolls that were surveyed all  years

(otherwise of coarse 2018 will have higher abundances...more knolls)

```{r}

knolls_2018<-fish_prep %>% filter(Year %in% c("2018")) %>% distinct(knoll_id)

knolls_2020_a<-fish_prep %>% filter(year_month %in% c("2020 a")) %>% distinct(knoll_id)

knolls_2020_b<-fish_prep %>% filter(year_month %in% c("2020 b")) %>% distinct(knoll_id)

knolls_2021_a<-fish_prep %>% filter(year_month %in% c("2021 a")) %>% distinct(knoll_id)



united_knolls <-Reduce(intersect,list(knolls_2018,knolls_2020_a,knolls_2020_b,knolls_2021_a))


fish_prep<-fish_prep %>% filter(knoll_id %in% united_knolls$knoll_id)

```

7. use all fish 

```{r}

all_fish <- fish_prep 

```

## prepre rank data

based on fuzzyQ of all fish from 2018 and on


```{r}

fuzzy_matrix <- all_fish %>% 
  select(Site,year_month,Knoll,Species,Abundance) %>%
  group_by(Site,year_month,Knoll,Species) %>% 
  summarise(Abundance=sum(Abundance))

fuzzy_matrix <- spread(fuzzy_matrix,
                       Species,
                       Abundance,
                       fill = 0)

fuzzy_matrix_species<-fuzzy_matrix[,4:ncol(fuzzy_matrix)]

fuzzy <- fuzzyq(fuzzy_matrix_species)

rank_data<-fuzzy[["spp"]]

rank_data$Species<-rownames(rank_data)

rownames(rank_data)<-NULL

rank_data<-rank_data %>%
  arrange(desc(Common.I)) %>%
  mutate("rank"=1:nrow(rank_data)) %>% 
  select(Species,rank,Common.I)


family_data<-all_fish %>% select(Family,Species) %>% distinct(Species,.keep_all = T)

rank_data<-left_join(rank_data,family_data)

```


recreate long data frame with all the species in all the knolls including 0 for species we didnt see

```{r}

ci_data<-fuzzy_matrix

ci_data_all<-gather(ci_data,key = "Species",value = "Abundance",4:ncol(ci_data))

```


# log ratios

## log ratio for 2018 vs strom data

```{r}
ci_data<-ci_data_all %>% arrange(Site,year_month,Knoll) %>% filter(year_month %in% c("2018 b","2020 a"))

calc_species_se = batch_calc_ES(dat = ci_data, grouping =  Species, condition = year_month, outcome = Abundance,scale = "count", ES = "LRRi")
```

prepre 2018 vs strom data

```{r}

log_data <- all_fish %>%
  filter(year_month %in% c("2018 b","2020 a")) %>% 
  group_by(year_month,Family,Species) %>%
  summarise(Abundance = sum(Abundance))

log_data <- spread(log_data,year_month,Abundance)

log_data <- left_join(rank_data,log_data)

log_data <- left_join(log_data,calc_species_se)


log_data <- log_data %>%
  rename("before" = `2018 b`,"after"= `2020 a`)



```

### plot 2018 vs storm

```{r}

log_data$Species_rank<-paste0(log_data$Species," (",log_data$rank,")")

log_data$Species <- reorder(log_data$Species, -log_data$Est)
log_data$Species_rank <- reorder(log_data$Species_rank, -log_data$Est)

log_data_short <- log_data %>% filter(rank < rank_filter)



ggplot(data = log_data_short,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  geom_errorbar(aes( xmin= CI_lower  , xmax= CI_upper),width = 0.5)+
  theme_classic()+xlim(-5,5)+
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =26,face = "italic"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.text = element_text(size =22,face = "italic"))+
  ggtitle("spring 2020/ Fall 2018")

#ggsave("log_ratio_storm_2018.pdf" , width = 15, height = 22, dpi = 300,path = plot_path)

cor.test(log_data$Common.I,log_data$Est)


```

## log ratio for 2018 vs fall 2020 (PostS2)
                
```{r}
ci_data_2<-ci_data_all %>% arrange(Site,year_month,Knoll) %>% filter(year_month %in% c("2018 b","2020 b"))

calc_species_se_2 = batch_calc_ES(dat = ci_data_2, grouping =  Species, condition = year_month, outcome = Abundance,scale = "count", ES = "LRRi")

```
                
    
```{r}

log_data_2 <- all_fish %>%
  filter(year_month %in% c("2018 b","2020 b")) %>% 
  group_by(year_month,Family,Species) %>%
  summarise(Abundance = sum(Abundance))

log_data_2 <- spread(log_data_2,year_month,Abundance)

log_data_2 <- left_join(rank_data,log_data_2)
log_data_2 <- left_join(log_data_2,calc_species_se_2)


log_data_2 <- log_data_2 %>%
  rename("before" = `2018 b`,"after"= `2020 b`) 


```

# plot

```{r}

log_data_2$Species <- reorder(log_data_2$Species, -log_data_2$Est)
log_data_2$Species_rank <- reorder(log_data_2$Species_rank, -log_data_2$Est)

log_data_short_2 <- log_data_2 %>% filter(rank < rank_filter)

ggplot(data = log_data_short_2,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  geom_errorbar(aes( xmin= CI_lower  , xmax= CI_upper),width = 0.5)+
  theme_classic()+
  scale_y_discrete(position = "right")+xlim(-5,5)+
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =26,face = "italic"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22))+
  ggtitle("Fall 2020/ Fall 2018")


#ggsave("log_ratio_F18_F20.pdf" , width = 15, height = 22, dpi = 300,path =plot_path)

cor.test(log_data_2$Common.I,log_data_2$Est)

```

### log ratios for 2020 spring (posts1) to 2021 spring (posts3)
                                                
         
                                                
```{r}
ci_data_3<-ci_data_all %>% arrange(Site,year_month,Knoll) %>% filter(year_month %in% c("2020 a","2021 a"))

calc_species_se_3 = batch_calc_ES(dat = ci_data_3, grouping =  Species, condition = year_month, outcome = Abundance,scale = "count", ES = "LRRi")
```
                                                
                                                
                                                
```{r}
log_data_3 <- all_fish %>%
  filter(year_month %in% c("2020 a","2021 a")) %>% 
  group_by(year_month,Family,Species) %>%
  summarise(Abundance = sum(Abundance))

log_data_3 <- spread(log_data_3,year_month,Abundance)

log_data_3 <- left_join(rank_data,log_data_3)
log_data_3 <- left_join(log_data_3,calc_species_se_3)
log_data_3 <- log_data_3 %>%
  rename("storm" = `2020 a`,"after"= `2021 a`) %>%
  mutate(ratio = after/storm, log_ratio = log(ratio))



```
                                                
                                                
# plot
                                                
```{r}


log_data_3$Species_rank<-paste0(log_data_3$Species," (",log_data_3$rank,")")

log_data_3$Species <- reorder(log_data_3$Species, -log_data_3$Est)
log_data_3$Species_rank <- reorder(log_data_3$Species_rank, -log_data_3$Est)

log_data_short_3 <- log_data_3 %>% filter(rank < rank_filter)

ggplot(data = log_data_short_3,aes(x=Est,y=Species_rank,fill = Family))+
  geom_bar(stat="identity",color = "black")+
  geom_errorbar(aes( xmin= CI_lower  , xmax= CI_upper),width = 0.5)+
  theme_classic()+xlim(-5,5)+
  theme(text = element_text(family = "serif"),
        axis.text.x = element_text(size=24),
        axis.text.y = element_text(size =26,face = "italic"),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22))+
  ggtitle("Spring 2020/ Spring 2021")


#ggsave("log_ratio_S20_S21.pdf" , width = 15, height = 22, dpi = 300,path =plot_path)

cor.test(log_data_3$Common.I,log_data_3$Est)

```
