---
title: "2.3 traits"
author: "TG"
date: "2023-07-18"
output:
   prettydoc::html_pretty:
    theme: cayman
    highlight: github
---


```{r message=FALSE, warning=FALSE}

library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(Rarefy)
library(rareNMtests)
library(zetadiv)
library(FuzzyQ)
library(SingleCaseES)
library(sjPlot)

```


# Data prepartion

```{r}
load("fish_full_data.rdata") # fish data

# traits data

traits_avery<-read.csv("traits data/traits_for_tal.csv")
Final_traits<-read.csv("traits data/Final_traits.csv")

select <- dplyr::select

fish_prep<-fish_full_data
```


# trait prepartion

Ive got 2 files sources:

1. avery table:

```{r}


my_species<-fish_full_data %>% distinct(Species)

traits_avery<-traits_avery %>% rename("Species"="species_fishbase_name")

my_species<-left_join(my_species,traits_avery)

trait_matrix<-my_species %>% select(Species,order,family,genus,
                                    home_range,activity,schooling,diet,water_level)

```

2. lab file

```{r}



Final_traits$Name<-str_to_sentence(Final_traits$Name)
size_traits<-Final_traits %>% select(Name,Size.FishBase,Size_type,Size.Class)
size_traits<-size_traits %>% rename("Species"="Name")

trait_matrix<-left_join(trait_matrix,size_traits)

trait_matrix %>% summarise_all(funs(sum(is.na(.))))

trait_matrix$Size.Class<-ordered(trait_matrix$Size.Class,c("S1","S2","S3","S4","S5","S6"))
trait_matrix$activity<-as.factor(trait_matrix$activity)
trait_matrix$diet<-as.factor(trait_matrix$diet)
trait_matrix$home_range<-as.factor(trait_matrix$home_range)
trait_matrix$water_level<-as.factor(trait_matrix$water_level)
trait_matrix$schooling<-as.factor(trait_matrix$schooling)


#save(trait_matrix,file ="ecocamp_trait_matrix.RData")

```


2. keep only data from 2018 
```{r}

fish_prep<- fish_prep %>% filter(Year > 2017)

```



3. remove irregular sites


```{r}

knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

fish_prep<-fish_prep %>% filter(!Knoll %in% knolls_to_remove) %>% 
  filter(Site !="North beach")

```

4. keep only sure identifications

filter confidence = 0


```{r}

fish_prep<-fish_prep %>% filter(Confidence <1)

```

5. keep one observer from each knoll


```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(fish_prep$survey_id)) {
  
  data <- fish_prep %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
  
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

fish_prep <- bind_rows(one_observer)

rm(data,filter_data,one_observer)

# make sure i have the right knolls number


knoll_site_year <-fish_prep %>% group_by(Site,year_month) %>% summarise("knoll_number"= n_distinct(knoll_id))

```


6. convert year_month to factor

```{r}

fish_prep$year_month<-as.factor(fish_prep$year_month)

```



7. keep only the knolls that were surveyed all  years

(otherwise of coarse 2018 will have higher abundances...more knolls)

```{r}

knolls_2018<-fish_prep %>% filter(Year %in% c("2018")) %>% distinct(knoll_id)

knolls_2020_a<-fish_prep %>% filter(year_month %in% c("2020 a")) %>% distinct(knoll_id)

knolls_2020_b<-fish_prep %>% filter(year_month %in% c("2020 b")) %>% distinct(knoll_id)

knolls_2021_a<-fish_prep %>% filter(year_month %in% c("2021 a")) %>% distinct(knoll_id)



united_knolls <-Reduce(intersect,list(knolls_2018,knolls_2020_a,knolls_2020_b,knolls_2021_a))


fish_prep<-fish_prep %>% filter(knoll_id %in% united_knolls$knoll_id)

```

7. use all fish 

```{r}

all_fish <- fish_prep 

```


## prepre rank data

based on fuzzyQ of all fish from 2018 and on


```{r}

fuzzy_matrix <- all_fish %>% 
  select(Site,year_month,Knoll,Species,Abundance) %>%
  group_by(Site,year_month,Knoll,Species) %>% 
  summarise(Abundance=sum(Abundance))

fuzzy_matrix <- spread(fuzzy_matrix,
                       Species,
                       Abundance,
                       fill = 0)

fuzzy_matrix_species<-fuzzy_matrix[,4:ncol(fuzzy_matrix)]

fuzzy <- fuzzyq(fuzzy_matrix_species)

rank_data<-fuzzy[["spp"]]

rank_data$Species<-rownames(rank_data)

rownames(rank_data)<-NULL

rank_data<-rank_data %>%
  arrange(desc(Common.I)) %>%
  mutate("rank"=1:nrow(rank_data)) %>% 
  select(Species,rank,Common.I)


family_data<-all_fish %>% select(Family,Species) %>% distinct(Species,.keep_all = T)

rank_data<-left_join(rank_data,family_data)

```


recreate long data frame with all the species in all the knolls including 0 for species we didnt see

```{r}

ci_data<-fuzzy_matrix

ci_data_all<-gather(ci_data,key = "Species",value = "Abundance",4:ncol(ci_data))

```


clean the environment

```{r}
rm(list=ls()[! ls() %in% c("ci_data_all","all_fish","fish_full_data","rank_data","trait_matrix")])


color_op6 <- c("#20A39E","#23001E","#FFBA49","#EF5B5B")
rank_filter <-51
plot_path<-"G:\\My Drive\\PHD\\chapter 1 strom effect\\R_storm_ms\\plots\\Log raito"
trait_path<-"G:\\My Drive\\PHD\\chapter 1 strom effect\\R_storm_ms\\plots\\traits"
model_path<-"G:\\My Drive\\PHD\\chapter 1 strom effect\\R_storm_ms\\models\\Traits"
```



# log ratios

## log ratio for 2018 vs strom data

```{r}
ci_data<-ci_data_all %>% arrange(Site,year_month,Knoll) %>% filter(year_month %in% c("2018 b","2020 a"))

calc_species_se = batch_calc_ES(dat = ci_data, grouping =  Species, condition = year_month, outcome = Abundance,scale = "count", ES = "LRRi")
```

prepre 2018 vs strom data

```{r}

log_data <- all_fish %>%
  filter(year_month %in% c("2018 b","2020 a")) %>% 
  group_by(year_month,Family,Species,Status) %>%
  summarise(Abundance = sum(Abundance))

log_data <- spread(log_data,year_month,Abundance)

log_data <- left_join(rank_data,log_data)

log_data <- left_join(log_data,calc_species_se)


log_data <- log_data %>%
  rename("before" = `2018 b`,"after"= `2020 a`)


rm(ci_data)
rm(calc_species_se)


```



### trait analyses



```{r}

trait_1<-left_join(log_data,trait_matrix)

trait_1<-trait_1 %>% drop_na(16:23)
  
filter_trait_1<-trait_1 %>% filter(rank<101)

hist(filter_trait_1$Est)


filter_trait_1$Status<-as.factor(filter_trait_1$Status)
filter_trait_1$Status[filter_trait_1$Species == "Parupeneus cyclostomus"]<-"T"

m1<-lmer(Est ~ diet+ Size.FishBase + water_level+schooling+rank+activity+ Status+(1|Family),data =filter_trait_1,na.action = "na.fail")
car::vif(m1)

dredge_table<-dredge(global.model = m1)

#write.csv(dredge_table,paste0(model_path,"\\traits_1_derdge_raw.csv"))

# best model:

dredge_table<-dredge_table[1:10,]

fm1<-lmer(Est ~Size.FishBase + water_level+ (1|Family),data =filter_trait_1,na.action = "na.fail")

summary(fm1)
plot(fm1)


```

```{r}
r1_list<-list()

#1

fm1<-lmer(Est ~Size.FishBase + (1|Family),data =filter_trait_1,na.action = "na.fail")

summary(fm1)

r1_list[[1]]<-as.data.frame(r.squaredGLMM(fm1))

#2

fm2<-lmer(Est ~Size.FishBase + water_level+ (1|Family),data =filter_trait_1,na.action = "na.fail")

summary(fm2)

r1_list[[2]]<-as.data.frame(r.squaredGLMM(fm2))

#3

fm3<-lmer(Est ~ + (1|Family),data =filter_trait_1,na.action = "na.fail")

summary(fm3)

r1_list[[3]]<-as.data.frame(r.squaredGLMM(fm3))

#4

fm4<-lmer(Est ~ water_level+ (1|Family),data =filter_trait_1,na.action = "na.fail")

summary(fm4)

r1_list[[4]]<-as.data.frame(r.squaredGLMM(fm4))


#5

fm5<-lmer(Est ~Size.FishBase + diet+ (1|Family),data =filter_trait_1,na.action = "na.fail")

summary(fm5)

r1_list[[5]]<-as.data.frame(r.squaredGLMM(fm5))
 
#6

fm6<-lmer(Est ~Status + (1|Family),data =filter_trait_1,na.action = "na.fail")

summary(fm6)

r1_list[[6]]<-as.data.frame(r.squaredGLMM(fm6))

#7

fm7<-lmer(Est ~Size.FishBase + Status+ (1|Family),data =filter_trait_1,na.action = "na.fail")

summary(fm7)

r1_list[[7]]<-as.data.frame(r.squaredGLMM(fm7))

#8

fm8<-lmer(Est ~water_level + Status+ (1|Family),data =filter_trait_1,na.action = "na.fail")

summary(fm8)

r1_list[[8]]<-as.data.frame(r.squaredGLMM(fm8))


#9

fm9<-lmer(Est ~water_level + activity+ (1|Family),data =filter_trait_1,na.action = "na.fail")

summary(fm9)

r1_list[[9]]<-as.data.frame(r.squaredGLMM(fm9))

#10

fm10<-lmer(Est ~Size.FishBase + activity+ (1|Family),data =filter_trait_1,na.action = "na.fail")

summary(fm10)

r1_list[[10]]<-as.data.frame(r.squaredGLMM(fm10))


r1_data<-bind_rows(r1_list)

dredge_table<-cbind(dredge_table,r1_data)

dredge_table$id<-rep("PostS1/PreS1")
#write.csv(dredge_table,paste0(model_path,"\\traits_1_derdge_short_r.csv"))


```







```{r}
#ggeffects::ggpredict(fm1, terms = c("Family"), type = "re") |> plot()
#ggeffects::ggpredict(fm1, terms = c("water_level")) |> plot()
#fm1_model_gg = ggeffects::ggpredict(fm1,terms = c('Size.FishBase[1:80 by=5]'))

colnames(fm1_model_gg)<-c("Size.FishBase","Est","std.error","conf.low","conf.high","group")
ggplot()+
  theme_classic()+
  geom_line(data = fm1_model_gg,aes(x= Size.FishBase,y = Est),size=1)+ 
  geom_ribbon(data = fm1_model_gg,aes(x=Size.FishBase,ymin=conf.low,ymax=conf.high),alpha=0.1)+
  geom_point(data = filter_trait_1,aes(x=Size.FishBase,y= Est,color = Family),size = 2)+
  ylab("Log ratio(PostS1/PreS1)")+
  xlab("Fish size (cm)")+
    theme(axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =20),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20),
        strip.text.x = element_text(size = 20))


#ggsave("posts1_pres1_model.png" , width = 12, height = 5, dpi = 300,path = trait_path)
#ggsave("posts1_pres1_model.pdf" , width = 12, height = 5, dpi = 300,path = trait_path)
```

### family model

```{r}
fm_fix<- lm(Est ~ Family,data =filter_trait_1,na.action = "na.fail")
summary(fm_fix)

```


# log ratios for 2020 spring (posts1) to 2021 spring (posts3)
                                                
         
                                                
```{r}
ci_data_3<-ci_data_all %>% arrange(Site,year_month,Knoll) %>% filter(year_month %in% c("2020 a","2021 a"))

calc_species_se_3 = batch_calc_ES(dat = ci_data_3, grouping =  Species, condition = year_month, outcome = Abundance,scale = "count", ES = "LRRi")

```
                                                
                                                
                                                
```{r}
log_data_3 <- all_fish %>%
  filter(year_month %in% c("2020 a","2021 a")) %>% 
  group_by(year_month,Family,Status,Species) %>%
  summarise(Abundance = sum(Abundance))

log_data_3 <- spread(log_data_3,year_month,Abundance)

log_data_3 <- left_join(rank_data,log_data_3)
log_data_3 <- left_join(log_data_3,calc_species_se_3)
log_data_3 <- log_data_3 %>% rename("storm" = `2020 a`,"after"= `2021 a`) 



```
                                            
                                                
```{r}
trait_3<-left_join(log_data_3,trait_matrix)

trait_3<-trait_3 %>% drop_na(16:23)

trait_3$Status<-as.factor(trait_3$Status)
  
filter_trait_3<-trait_3 %>% filter(rank<101)

hist(filter_trait_3$Est)

m3_1<-lmer(Est ~ diet+ Size.FishBase + water_level  +schooling+rank+activity+Status+ (1|Family),data =filter_trait_3,na.action = "na.fail")


dredge_table_3<-dredge(global.model = m3_1)

#write.csv(dredge_table_3,paste0(model_path,"\\traits_3_derdge_raw.csv"))

dredge_table_3<-dredge_table_3[1:10,]



#fm3<-lmer(Est ~ water_level + (1|Family),data =filter_trait_3,na.action = "na.fail")
#summary(fm3)
#ggeffects::ggpredict(fm3, terms = c("water_level")) |> plot()
#ggeffects::ggpredict(fm3, terms = c("Family"), type ="re") |> plot()
#fm1_model_gg = ggeffects::ggpredict(fm1,terms = c('Size.FishBase[1:80 by=5]'))




```

```{r}


rm_list<-list()

#1

rm1<-lmer(Est ~ water_level + (1|Family),data =filter_trait_3,na.action = "na.fail")
summary(rm1)
rm_list[[1]]<-as.data.frame(r.squaredGLMM(rm1))

#2

rm2<-lmer(Est ~  (1|Family),data =filter_trait_3,na.action = "na.fail")
summary(rm2)
rm_list[[2]]<-as.data.frame(r.squaredGLMM(rm2))


#3

rm3<-lmer(Est ~ Status+ (1|Family),data =filter_trait_3,na.action = "na.fail")
summary(rm3)
rm_list[[3]]<-as.data.frame(r.squaredGLMM(rm3))

#4

rm4<-lmer(Est ~ Status+ water_level+ (1|Family),data =filter_trait_3,na.action = "na.fail")
summary(rm4)
rm_list[[4]]<-as.data.frame(r.squaredGLMM(rm4))

#5

rm5<-lmer(Est ~ activity + (1|Family),data =filter_trait_3,na.action = "na.fail")
summary(rm5)
rm_list[[5]]<-as.data.frame(r.squaredGLMM(rm5))

#6

rm6<-lmer(Est ~ activity + water_level+(1|Family),data =filter_trait_3,na.action = "na.fail")
summary(rm6)
rm_list[[6]]<-as.data.frame(r.squaredGLMM(rm6))

#7

rm7<-lmer(Est ~ schooling + water_level+(1|Family),data =filter_trait_3,na.action = "na.fail")
summary(rm7)
rm_list[[7]]<-as.data.frame(r.squaredGLMM(rm7))

#8

rm8<-lmer(Est ~ schooling +activity+ water_level+(1|Family),data =filter_trait_3,na.action = "na.fail")
summary(rm8)
rm_list[[8]]<-as.data.frame(r.squaredGLMM(rm8))

#9

rm9<-lmer(Est ~ activity +Status+(1|Family),data =filter_trait_3,na.action = "na.fail")
summary(rm9)
rm_list[[9]]<-as.data.frame(r.squaredGLMM(rm9))

#10


rm10<-lmer(Est ~ schooling+(1|Family),data =filter_trait_3,na.action = "na.fail")
summary(rm10)
rm_list[[10]]<-as.data.frame(r.squaredGLMM(rm10))

rm_data<-bind_rows(rm_list)


dredge_table_3<-cbind(dredge_table_3,rm_data)


dredge_table_3$id<-rep("PostS3/PostS1")
```

```{r}
united_dredge<-rbind(dredge_table,dredge_table_3)

#write.csv(united_dredge,paste0(model_path,"\\united_dredge.csv"))
```


##### Family model 

```{r}
fm_fix_recovery<-lm(Est ~ Family,data =filter_trait_3,na.action = "na.fail")
summary(fm_fix_recovery)
plot(fm3)
```

print table for both family models

```{r}
#tab_model(fm_fix,fm_fix_recovery,file = paste0(model_path,"//fm_fix_1.doc"))
```

