---
title: "5.1 knoll trajectories"
author: "TG"
date: "2023-07-01"
output: html_document
---


```{r message=FALSE, warning=FALSE}

library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)
library(stringr)
library(Hmisc)
library(vegan)
library(mobr)
library(lme4)
library(lmerTest)
library(MuMIn)
library(plotrix)
library(forcats)


color_op6_l <- c("#B66DAF","#7CE171","#20A39E","#23001E","#FFBA49","#EF5B5B")

save_path_abundance<-"G:\\My Drive\\PHD\\chapter 1 strom effect\\R_storm_ms\\plots\\knoll trajectories\\abundance"
save_path_richness<-"G:\\My Drive\\PHD\\chapter 1 strom effect\\R_storm_ms\\plots\\knoll trajectories\\richness"

```


In this script ill concentrate all the models and plot constructing Fig 2

# Data prepartion

```{r}
load("fish_full_data.rdata")
select <- dplyr::select

fish_prep<-fish_full_data
```

## general steps:

1. keep all years exept 2017

```{r}

fish_prep <- fish_full_data %>% filter(Year!=2017)

```


2. remove irregular sites

```{r}
knolls_to_remove<-c("South bridge reserve","North bridge reserve","south bridge reserve",
                    "Tamar","north bridge reserve")

fish_prep<-fish_prep %>% filter(!Knoll %in% knolls_to_remove) %>% filter(Site !="North beach")
```


3. keep one observer from each knoll


```{r}
one_observer<-list()
loop<-1
set.seed(1)

for (i in unique(fish_prep$survey_id)) {
  
  data <- fish_prep %>% filter(survey_id == i)
  
  obs_amount <- length(unique(data$Observer))
  obs_name <- unique(data$Observer)
  chosen_obs <- sample(obs_name,1)  
 
  filter_data<-data %>% filter(Observer == chosen_obs)
  one_observer[[loop]]<-filter_data
  loop<-loop+1
  
  
}

fish_prep <- bind_rows(one_observer)

rm(data,filter_data,one_observer)

# make sure i have the right knolls number


knoll_site_year <-fish_prep %>% group_by(Site,year_month) %>% summarise("knoll_number"= n_distinct(knoll_id))

```

4. keep only the resident fish

```{r}

fish_prep <- fish_prep %>% filter(Status != "T")

```


5. convert year_month to factor

```{r}

fish_prep$year_month<-as.factor(fish_prep$year_month)
```


6. create the data for species richness, abundance and shannon model

General stepes:
1. add confisance to data before 2018

2. filter accourding to confidance level (richness and eveness; confidance = 0, abundance; confidance = 2)

3. filter schools for eveness

4. group by the right variable (number of species/ind/effective number)


```{r}
#early years didnt have confidance 

fish_prep$Confidence[is.na(fish_prep$Confidence)]<-(0)
```


### richness

```{r}

richness <- fish_prep %>% filter(Confidence < 1)

richness<- richness %>%
  group_by(survey_id) %>%
  mutate(Species_number = n_distinct(Species_2015)) %>%
  select(-Species,-Species_2015,-Family,-Abundance,-Length,-Confidence,-Status,-a,-b) %>%
  ungroup() %>%
  distinct(.keep_all = T)

```


```{r}

richness_bg<-richness %>% filter(Year <2020)
richness_aov<-aov(Species_number ~ knoll_id,data = richness_bg)
summary(richness_aov)


a<-richness_aov[["coefficients"]]
mean(richness_aov[["coefficients"]])
mean(richness_aov[["residuals"]])

lmer(Species_number ~ Site+(1|knoll_id),data = richness_bg)


#ggplot(data = richness_bg,aes(x=fct_reorder(Knoll,Species_number,.fun='max'),y=Species_number))+geom_boxplot()+facet_wrap(~Site)
```




### abundance

```{r}
abundance <- fish_prep %>%
  group_by(survey_id,Site,knoll_id,Year,year_month,Max_depth,Min_depth,Mean_Complexity,
           Surface_Area,Algae,coral_cover) %>%
    summarise("ind"=sum(Abundance))

which(!abundance$survey_id %in% richness$survey_id)


abundance_bg<-abundance %>% filter(Year <2020)
abundance_aov<-aov(ind ~ knoll_id,data = abundance_bg)

summary(abundance_aov)




school_species <- c("Chromis viridis",
                    "Pseudanthias spp.",
                    "Neopomacentrus miryae",
                    "Atherinidae",
                    "Spratelloididae") 

abundance_ns <- fish_prep %>% filter(!Species %in% school_species) %>% 
  group_by(survey_id,Site,knoll_id,Year,year_month,Max_depth,Min_depth,Mean_Complexity,
           Surface_Area,Algae,coral_cover) %>%
    summarise("ind"=sum(Abundance))

which(!abundance_ns$survey_id %in% richness$survey_id)


abundance_bg_ns<-abundance_ns %>% filter(Year <2020)
abundance_aov_ns<-aov(ind ~ knoll_id,data = abundance_bg_ns)

summary(abundance_aov_ns)

```

### Shannon

first - filter and convert to wide data format

chand to the school fish name that valid for 2015

Pseudanthias squamipinnis == Pseudanthias spp.

Atherinomorus lacunosus == Atherinidae

Spratelloides gracilis == Spratelloididae

```{r}

school_species <- c("Chromis viridis",
                    "Pseudanthias spp.",
                    "Neopomacentrus miryae",
                    "Atherinidae",
                    "Spratelloididae") 

shannon <- fish_prep %>% filter(! Species_2015 %in% school_species) %>% filter(Confidence < 1)


shannon <- shannon %>%
  group_by(Site,Year,year_month,Knoll,DD_X,DD_Y,Surface_Area,Max_depth,Date,survey_id,Species_2015) %>%
  summarise("aboundance"=sum(Abundance))


shannon<-spread(shannon,Species_2015,aboundance,fill = 0)

which(!abundance$survey_id %in% shannon$survey_id) # need to be 1, transect "Caves_11_2020-06-14" had only schooling core species

```
calculate the shannon effective number

```{r}
shannon$shannon_div<-diversity(shannon[,11:ncol(shannon)],index="shannon")

shannon<-shannon %>%
  relocate(shannon_div,.after=survey_id) %>% 
  mutate("shannon_effective"=exp(shannon_div),.after=shannon_div)
  


```

# Abundance plot for each knoll

```{r}


for (i in unique(abundance$Site)) {
  
  
p<-abundance %>% filter(Site == i) %>% 
  mutate(Knoll = as.factor(Knoll)) %>% 
  mutate(Knoll = fct_reorder(Knoll,ind,.fun='max')) %>%
  ggplot( aes(x=fct_reorder(Knoll,ind,.fun='max'), y=ind, color = year_month)) + geom_point(size=2,position=position_jitter(width=0, height=0.4)) +
  ggtitle(i)+
  xlab("Knoll id") +
  ylab("Resident fish abundance")+
 theme_bw()+
  theme(title = element_text(size = 20),
       axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values=color_op6_l,
                           name = "Year",
                           labels = c("Fall 2015","Fall 2016","Fall 2018","Spring 2020*", "Fall 2020" ,
                                      "Spring 2021"))

print(p)


p_pdf<-paste0("fish_abundance_",i,".pdf")
p_png<-paste0("fish_abundance_",i,".png")

ggsave(p_pdf, width = 14, height =5, dpi = 300,path = save_path_abundance)
ggsave(p_png, width = 14, height =5, dpi = 300,path = save_path_abundance)   
}
  



```



```{r}
for (i in unique(richness$Site)) {
  
  
p<-richness %>% filter(Site == i) %>% 
  mutate(Knoll = as.factor(Knoll)) %>% 
  mutate(Knoll = fct_reorder(Knoll,Species_number,.fun='max')) %>%
  ggplot( aes(x=fct_reorder(Knoll,Species_number,.fun='max'), y=Species_number, color = year_month)) + geom_point(size=2,position=position_jitter(width=0, height=0.4)) +
  ggtitle(i)+
  xlab("Knoll id") +
  ylab("Resident fish richness")+
 theme_bw()+
  theme(title = element_text(size = 20),
       axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values=color_op6_l,
                           name = "Year",
                           labels = c("Fall 2015","Fall 2016","Fall 2018","Spring 2020*", "Fall 2020" ,
                                      "Spring 2021"))

print(p)


p_pdf<-paste0("fish_richness_",i,".pdf")
p_png<-paste0("fish_richness_",i,".png")

ggsave(p_pdf, width = 14, height =5, dpi = 300,path = save_path_richness)
ggsave(p_png, width = 14, height =5, dpi = 300,path = save_path_richness)   
}
  

```

```{r}
abundance %>% filter(Site =="NR") %>% 
  mutate(Knoll = as.factor(Knoll)) %>% 
  mutate(Knoll = fct_reorder(Knoll,ind,.fun='max')) %>%
  ggplot( aes(x=fct_reorder(Knoll,ind,.fun='max'), y=ind, color = year_month)) + geom_point(size=2,position=position_jitter(width=0, height=0.4)) +
  ggtitle("NR")+
  xlab("Knoll id") +
  ylab("Resident fish abumdance")+
 theme_bw()+
  theme(title = element_text(size = 20),
       axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))+
  scale_color_manual(values=color_op6_l,
                           name = "Year",
                           labels = c("Fall 2015","Fall 2016","Fall 2018","Spring 2020*", "Fall 2020" ,
                                      "Spring 2021"))




abundance %>% filter(Site =="NR") %>% 
  ggplot( aes(x=year_month, y=ind, color = Knoll,group = Knoll)) + 
  geom_point(size=2,position=position_jitter(width=0, height=0.4)) +
  geom_path()+
  ggtitle("NR")+
  xlab("Knoll id") +
  ylab("Resident fish abumdance")+
 theme_bw()+
  theme(title = element_text(size = 20),
       axis.text.x = element_text(size=20),
        axis.text.y = element_text(size =16),
        axis.title.x = element_text(size=22),
        axis.title.y = element_text(size=22),
        legend.title = element_text(size=22),
        legend.text = element_text(size=20))

+
  scale_color_manual(values=color_op6_l)+
   labs(color = "Year")
```

